Environmental and geographic variables are effective surrogates for genetic variation in conservation planning

### 2016-08-22 code/rmarkdown/article.Rmd
+ use checkpoint in properly
+ upgrade article deps to 3.3.1
+ assorted article fixes

```{r,include=FALSE}
# load .rda file
try(session::restore.session('data/final/results.rda'))

@@

First, we found that on average the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$). This result indicates that regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive then neutral genetic variation.

@@

Second, the proportion of genetic variation secured in a prioritisation was found to depend on the combination of targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$. Specifically, the amount-based prioritisations (`r round(mean(scenario.sub.DF,filter(Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; `$Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$). No evidence was found to suggest that the prioritisations that included additional surrogate-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100` %) or genetic-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$). These results suggest that the addition of environmental and geographic surrogates can improve the proportion of intra-specific genetic variation secured in a prioritisation for most species.

@@

Third, we also found significant differences in the average proportion of genetic variation secured by prioritisations generated under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`). Specifically, both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were both found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively). Neither of the multi-species scenarios were found to secure a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$). These results suggest that on average the multi-species prioritisations can secure more genetic variation than single-species prioritisations--regardless of targets used to generate them.

@@

<!-- suitable targets for surrogates -->
Generally, the slopes correlating the proportion of geographic variation and neutral variation secured in a prioritisation were higher than those for the proportion of environmental variation and adaptive variation secured in a prioritisation (Figures 2a--2b and S56--S57). To secure a representative sample of each species' adaptive genetic variation ($>=$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), the prioritisation would need to sample on average `r mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` % $\pm$ `r sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` SD of the environmental conditions found across the planning units occupied by the species. On the other hand, to secure a representative sample of the species' neutral genetic variation ($>=$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100`), the prioritisation would need to sample % `r mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` $\pm$ `r sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` SD of the variation in geographic position among the planning units occupied by the species. These results suggest that geographic surrogates may require higher targets than environmental surrogates to generate prioritisations that secure representative sample of genetic variation. Additionally, these results suggest using surrogate-based targets alone to generate prioritisations may be insufficient to secure a representative sample of species' genetic variation. For all species, prioritisations were generated that secured a high proportion of the surrogate variable (> 90 %) but also secured a minuscule proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57). Thus we recommend that prioritisations be generated using a combination of amount-based targets--as is conventionally used in reserve selection processes--and also surrogate-based targets. After establishing a relationship between the proportion of surrogate variation and genetic variation secured in prioritisation, we than generated prioritisations under various planning scenarios to provide insight into how the surrogates may work in the real-world.

@@

<!-- neutral genetic variation secured in prioritisations-->
Generally, the addition of environmental and geographic targets were needed to obtain prioritisations that secure a representative proportion of neutral genetic variation (Figure 4). With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($>=$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h). Whereas, in all scenarios, the prioritisations generated using amount and surrogate based targets secured a representative proportion of neutral genetic variation for most species. Although the the latter prioritisation did not secure a representative sample of the neutral genetic variation in all species, this result suggests that the use of environmental and geographic targets in conservation planning can result in a more effective prioritisation. Furthermore, this improvement was associated with only a moderate increase in cost in both scenarios (scenario without opportunity: `r round((score(multi.spp.prioritisations[[2]])/score(multi.spp.prioritisations[[1]]))*100,2)` %; scenario with opportunity cost: `r  round((score(multi.spp.prioritisations.with.cost[[2]])/score(multi.spp.prioritisations.with.cost[[1]]))*100,2)` %)--emphasising the utility of surrogates in conservation.

>>>>

```{r,include=FALSE}
# load .rda file
checkpoint::checkpoint('2016-08-03', R.version='3.3.1', scanForPackages=FALSE)
try(session::restore.session('data/final/results.rda'))

@@

First, we found that on average the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`). This result indicates that regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive then neutral genetic variation.

@@

Second, the proportion of genetic variation secured in a prioritisation was found to depend on the combination of targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$. Specifically, the amount-based prioritisations (`r round(mean(scenario.sub.DF,filter(Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$). No evidence was found to suggest that the prioritisations that included additional surrogate-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100` %) or genetic-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$). These results suggest that the addition of environmental and geographic surrogates can improve the proportion of intra-specific genetic variation secured in a prioritisation for most species.

@@

Third, we also found significant differences in the average proportion of genetic variation secured by prioritisations generated under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`). Specifically, both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were both found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively). Neither of the multi-species scenarios were found to secure a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$). These results suggest that on average the multi-species prioritisations can secure more genetic variation than single-species prioritisations--regardless of targets used to generate them.

@@

<!-- suitable targets for surrogates -->
Generally, the slopes correlating the proportion of geographic variation and neutral variation secured in a prioritisation were higher than those for the proportion of environmental variation and adaptive variation secured in a prioritisation (Figures 2a--2b and S56--S57). To secure a representative sample of each species' adaptive genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), the prioritisation would need to sample on average `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental conditions found across the planning units occupied by the species. On the other hand, to secure a representative sample of the species' neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100`), the prioritisation would need to sample % `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species. These results suggest that geographic surrogates may require higher targets than environmental surrogates to generate prioritisations that secure representative sample of genetic variation. Additionally, these results suggest using surrogate-based targets alone to generate prioritisations may be insufficient to secure a representative sample of species' genetic variation. For all species, prioritisations were generated that secured a high proportion of the surrogate variable (> 90 %) but also secured a minuscule proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57). Thus we recommend that prioritisations be generated using a combination of amount-based targets--as is conventionally used in reserve selection processes--and also surrogate-based targets. After establishing a relationship between the proportion of surrogate variation and genetic variation secured in prioritisation, we than generated prioritisations under various planning scenarios to provide insight into how the surrogates may work in the real-world.

@@

<!-- neutral genetic variation secured in prioritisations-->
Generally, the addition of environmental and geographic targets were needed to obtain prioritisations that secure a representative proportion of neutral genetic variation (Figure 4). With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h). Whereas, in all scenarios, the prioritisations generated using amount and surrogate based targets secured a representative proportion of neutral genetic variation for most species. Although the the latter prioritisation did not secure a representative sample of the neutral genetic variation in all species, this result suggests that the use of environmental and geographic targets in conservation planning can result in a more effective prioritisation. Furthermore, this improvement was associated with only a moderate increase in cost in both scenarios (scenario without opportunity: `r round((score(multi.spp.prioritisations[[2]])/score(multi.spp.prioritisations[[1]]))*100,2)` %; scenario with opportunity cost: `r  round((score(multi.spp.prioritisations.with.cost[[2]])/score(multi.spp.prioritisations.with.cost[[1]]))*100,2)` %)--emphasising the utility of surrogates in conservation.
